\textbf{Advection versus diffusion in intraventricular transport}

### 2025-03-25 paper/main.tex
minor fixes

Ciliary motion is integral to physiological CSF flow,
and absence of cilia has been associated with defective
brain development and various pathologies~\cite{Eichele2020Cilia-drivenVentricle,
Olstad2019CiliaryDevelopment, Faubel2016Cilia-basedVentricles,
Guirao2010CouplingCilia, Afzelius2004CiliaRelatedDiseases,
Sawamoto2006NewBrain, Yoshiba2014RolesSymmetry,
Hirokawa2006NodalAsymmetry}.
As observed both experimentally and computationally,
the absence of motile cilia alters CSF flow patterns. 
We next investigate the implications of these changes on intraventricular
transport.

@@

A possible improvement to the in- and outflow boundary conditions would be
to consider a 3D-0D coupling of the in- and outflow boundaries using
ordinary differential equations,
similar to what has been done in cardiac
modeling~\cite{Brown2024AMechanics, Augustin2021ACirculation}.
However, if the ventricular system in reality is a closed system,
one should rather seek out a way of modeling the CSF flow without open boundaries to avoid
handling in- and outflow boundary conditions. A more appropriate way of handling the beating
cardiac motion could be to impose a ventricular wall deformation, assumed to result from pulsatile
motion of the heart and/or cardiac vessels surrounding the ventricles. This wall deformation would
induce CSF flow while simultaneously permitting enforcement of the impermeability condition
$\uu\cdot\nn=0$ on the ventricular walls. Such wall-motion induced CSF flow has been modeled and
simulated successfully in the human brain ventricular system~\cite{Causemann2022HumanFramework,
Kurtcuoglu2005ComputationalSystem, Kurtcuoglu2007ComputationalSylvius, Linninger2005PulsatileBrain}.

The diffusion coefficients used in this study were based on experimental values. 
The value $D_3=\num{1.15e-10} \ \mathrm{m^2/s}$
is based on the value $\num{1.15e-10} \pm 0.11 \ \mathrm{m^2/s}$ reported in an
experiment~\cite{GuraSadovsky2017MeasurementExpansion}, while the Stokes-Einstein
relation~(\Cref{eq:D_stokes_einstein})~\cite{Einstein1905UberTeilchen} was used to extrapolate
the value of $D_2$ and estimate the value of $D_1$, based on experimental
data~\cite{Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells,
Mahmood2023ExosomeTemperature}. The Stokes-Einstein relation assumes the diffusing
substance is of spherical shape, which proteins are not, and the relation does not
necessarily hold in CSF. Moreover, the extrapolation implies the assumption that
the two substances (the one for which we determine $D$ and the one we extrapolate from)
have equal densities. Consequently, there is uncertainty in the diffusion coefficient values used.
We note that Sadovsky~\emph{et al.}~\cite{GuraSadovsky2017MeasurementExpansion} reported that
Dendra2 satisfies the Stokes-Einstein in a solution of similar viscosity
to CSF. We used CSF density and viscosity values measured in
humans~\cite{Bloomfield1998EffectsFluid},
and because embryonic CSF in zebrafish is high in proteins, nutrients and other substances,
the CSF in which the photoconverted Dendra2 are transported in our experiments may be more viscous
than we assumed. This would reduce the diffusion coefficient (cf.~\Cref{eq:D_stokes_einstein}),
but based on reports of protein concentration having a limited impact on the viscosity of
CSF~\cite{Bloomfield1998EffectsFluid, Brydon1995PhysicalViscosity}, this reduction is presumably
limited. Merely the presence of other proteins and nutrients may also reduce the diffusion
coefficient by reduction of mean-free path lengths of molecules~\cite{Cussler2009Diffusion:Systems}.
Considering the diffusion coefficients $D_1$ and $D_2$ (both smaller
than $D_3$) in our transport analysis provides information about how the simulation
results of Dendra2 transport would be altered if the
diffusion coefficient $D_3$ is overestimated. In the case that $D_3$ is underestimated, 
we expect a trend similar to the difference in the $D_2$ and $D_3$ solute concentrations;
incerasingly diffusive fields.

@@

In large brains such as the human brain, we presume ciliary beating mostly 
contributes to flow in localized regions proximal to the ventricular walls due to
the microscopic length of the cilia.
Although the precise mechanisms of intraventricular flow are still debated,
much evidence indicate that blood vessel pulsations, respiration, CSF secretion
and brain motion induce the pulsatile bulk flow in the central
regions of the ventricles~\cite{Vinje2019RespiratoryMeasurements,
Siyahhan2014FlowVentricles, Linninger2005PulsatileBrain, Enzmann1992BrainImaging,
Kurtcuoglu2007ComputationalSylvius}.
In human brains, cilia therefore probably play an important role in regulating the
amount of substances like extracellular vesicles close to the ventricular walls, similar to coral
reefs, where it has been observed that cilia are primary regulators of
oxygen amounts at the coral surface~\cite{Pacherres2022CiliaryProduction}.
Flow mediated by cilia indeed serves such a purpose in nodal flow at early
developmental stages of the brain~\cite{Hirokawa2006NodalAsymmetry}, and the discovery
of a cilia-driven flow network in rodent brains suggest such selective
transport~\cite{Faubel2016Cilia-basedVentricles}.

>>>>

Ciliary motion is integral to physiological CSF flow,
and absence of cilia has been associated with defective
brain development and various pathologies~\cite{Eichele2020Cilia-drivenVentricle,
Olstad2019CiliaryDevelopment, Faubel2016Cilia-basedVentricles,
Guirao2010CouplingCilia, Afzelius2004CiliaRelatedDiseases,
Sawamoto2006NewBrain, Yoshiba2014RolesSymmetry,
Hirokawa2006NodalAsymmetry}.
As observed both experimentally and computationally,
absence of motile cilia alters CSF flow patterns. 
We next investigate the implications of these changes on intraventricular
transport.

@@

A possible improvement to the in- and outflow boundary conditions would be
to consider a 3D-0D coupling of the in- and outflow boundaries using
ordinary differential equations,
similar to what has been done in cardiac
modeling~\cite{Brown2024AMechanics, Augustin2021ACirculation}.
However, if the ventricular system in reality is a closed system,
one should rather seek out a way of modeling the CSF flow without open boundaries to avoid
handling in- and outflow boundary conditions. A more appropriate way of handling the beating
cardiac motion could be to impose a ventricular wall deformation, assumed to result from pulsatile
motion of the heart and/or cardiac vessels surrounding the ventricles. This wall deformation would
induce CSF flow while simultaneously permitting enforcement of the impermeability condition
$\uu\cdot\nn=0$ on the ventricular walls. Such wall motion-induced CSF flow has been modeled and
simulated successfully in the human brain ventricular system~\cite{Causemann2022HumanFramework,
Kurtcuoglu2005ComputationalSystem, Kurtcuoglu2007ComputationalSylvius, Linninger2005PulsatileBrain}.

The diffusion coefficients used in this study are based on experimental values. 
The value $D_3=\num{1.15e-10} \ \mathrm{m^2/s}$
is based on the value $\num{1.15e-10} \pm 0.11 \ \mathrm{m^2/s}$ reported in an
experiment~\cite{GuraSadovsky2017MeasurementExpansion}, while the Stokes-Einstein
relation~(\Cref{eq:D_stokes_einstein})~\cite{Einstein1905UberTeilchen} was used to extrapolate
the value of $D_2$ and estimate the value of $D_1$, based on experimental
data~\cite{Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells,
Mahmood2023ExosomeTemperature}. The Stokes-Einstein relation assumes the diffusing
substance is of spherical shape, which proteins are not, and the relation does not
necessarily hold in CSF. Moreover, the extrapolation implies the assumption that
the two substances (the one for which we determine $D$ and the one we extrapolate from)
have equal densities. Consequently, there is uncertainty in the diffusion coefficient values used.
We note that Sadovsky~\emph{et al.}~\cite{GuraSadovsky2017MeasurementExpansion} reported that
Dendra2 satisfies the Stokes-Einstein in a solution of similar viscosity
to CSF. We used CSF density and viscosity values measured in
humans~\cite{Bloomfield1998EffectsFluid},
and because embryonic CSF in zebrafish is high in proteins, nutrients and other substances,
the CSF in which the photoconverted Dendra2 are transported in our experiments may be more viscous
than we assumed. This could reduce the diffusion coefficient (cf.~\Cref{eq:D_stokes_einstein}),
but based on reports of protein concentration having a limited impact on the viscosity of
CSF~\cite{Bloomfield1998EffectsFluid, Brydon1995PhysicalViscosity}, this reduction is presumably
limited. Merely the presence of other proteins and nutrients may also reduce the diffusion
coefficient by reduction of mean-free path lengths of molecules~\cite{Cussler2009Diffusion:Systems}.
Considering the diffusion coefficients $D_1$ and $D_2$ (both smaller
than $D_3$) in our transport analysis provides information about how the simulation
results of Dendra2 transport would be altered if the
diffusion coefficient $D_3$ is overestimated. In the case that $D_3$ is underestimated, 
we expect a trend similar to the difference in the $D_2$ and $D_3$ solute concentrations;
incerasingly diffusive fields.

@@

In large brains such as the human brain, we presume ciliary beating mostly 
contributes to flow in localized regions proximal to the ventricular walls, due to
the microscopic length of the cilia.
Although the precise mechanisms of intraventricular flow are still debated,
much evidence indicate that blood vessel pulsations, respiration, CSF secretion
and brain motion induce the pulsatile bulk flow in the central
regions of the ventricles~\cite{Vinje2019RespiratoryMeasurements,
Siyahhan2014FlowVentricles, Linninger2005PulsatileBrain, Enzmann1992BrainImaging,
Kurtcuoglu2007ComputationalSylvius}.
In human brains, cilia therefore probably play an important role in regulating the
amount of substances like extracellular vesicles close to the ventricular walls, similar to coral
reefs, where it has been observed that cilia are primary regulators of
oxygen amounts at the coral surface~\cite{Pacherres2022CiliaryProduction}.
Flow mediated by cilia indeed serves such a purpose in nodal flow at early
developmental stages of the brain~\cite{Hirokawa2006NodalAsymmetry}, and the discovery
of a cilia-driven flow network in rodent brains suggests such selective
transport~\cite{Faubel2016Cilia-basedVentricles}.
